---
title: "NEAIR Workshop"
author: ""
date: ""
output: html_document
---

This R markdown file is used to teach R for Institutional Researchers.

```{r setup, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
# This R chunk will be used to prepare the data for analysis. Typically, we will
# want to hide the R code and output in the final document.

# Load the packages we will use
library(ggplot2)
library(psych)
library(readxl)
library(knitr)

# Load data

# Convert Military variable to logical

# Convert Enrolled variable to a Date

```

There are `r ` in the `students` data set.

## Descriptive Statistics

The following tables summarizes the number of credits students have by Division.

```{r credits, echo=TRUE, results='asis'}

```

## Number of Military Students by Division

```{r militaryTable, echo=TRUE}

# Print table as counts

# Print table as proportion within division

```

#### Barplot of military students by division

```{r militaryBarplot, echo=TRUE}
# Hint: use position='dodge' to create side-by-side bars

```

## Distribution of Credits

```{r histogram, echo=TRUE}
# Use histogram

```

## Distribution of Time Enrolled by Division

```{r enrolled, echo=TRUE}
# Use boxplots

```

## Is there a difference in the number of credits by military and civilian students?

```{r miltaryCredits, echo=TRUE}
# What to do about NA credits? In this case missing credits mean 0

# t-test maybe?


# Create a function to print the results of p-value as p < 0.001, p < 0.01, p < 0.05 or p > 0.05

```

The difference between military and civilian credits is `r ` which is `r ` statistically significant with `r `

