---
title: "Intro to R Workshop"
output: html_document
---

# Import

In the R chunk below, load in the `students.csv`, `coursework.csv`, and `demographics.csv` files into R.

```{r import-read}
students <- read.csv('Data/students.csv', stringsAsFactors = FALSE)
coursework <- read.csv('Data/coursework.csv', stringsAsFactors = FALSE)
demographics <- read.csv('Data/demographics.csv', stringsAsFactors = FALSE)
```

## Viewing Data

Print the first six rows of `students` and the last six rows of `coursework`.

```{r import-head}
head(students)
tail(coursework)
```

What is the `str`ucture of these data frames?

```{r import-str}
str(students)
str(coursework)
str(demographics)
```

## Subsetting Data

Print all the students who live in Vermont.

```{r import-select}
students[which(students$STATE_CODE == 'VT'),]
```

Whare are all the different grade codes students have.

```{r import-unique}
unique(coursework$GRADE_CODE)
```


# Tidy

Convert the appropriate date columns to actual dates (if they are not already).

```{r tidy-convertdates}
students$ENROLL_DATE <- as.Date(students$ENROLL_DATE)
students$DATE_OF_BIRTH <- as.Date(students$DATE_OF_BIRTH)
coursework$DATE_TAKEN <- as.Date(coursework$DATE_TAKEN)
```

Convert `GENDER` to a factor.

```{r tidy-factor}
students$GENDER <- as.factor(students$GENDER)
```

Convert `GRADE_CODE` to an ordered factor. You may wish to first convert all withdrawn grades (AW, MW, and W) to missing values first (i.e. `NA`).

```{r tidy-orderedfactor}
coursework[coursework$GRADE_CODE %in% c('AW','MW','W','I'),]$GRADE_CODE <- NA
unique(coursework$GRADE_CODE)
coursework$GRADE_CODE <- factor(coursework$GRADE_CODE,
					            levels = c('F','D','C','B','P','A'),
					            ordered = TRUE)
```

Convert any missing values for `TRANSFER_CREDITS_EARNED` to zero.

```{r tidy-missingtozero}
students[is.na(students$TRANSFER_CREDITS_EARNED),]$TRANSFER_CREDITS_EARNED <- 0
```


# Transform

What are the number of students for each ethnicity group?

```{r transform-table}
table(students$ETHNICITY, useNA='ifany')
```

What is the proportion of students in each ethnicity group?

```{r transform-proptable}
prop.table(
	table(students$ETHNICITY, useNA='ifany')
) * 100
```

What is the number of students in each ethnicity group by gender?

```{r transform-table2}
table(students$ETHNICITY, students$GENDER, useNA='ifany')
```

What is the proportion of students in each ethnicity group by gender where the rows equal 1 (or 100)?

```{r transform-proptable2}
prop.table(	table(students$ETHNICITY, students$GENDER, useNA='ifany'), 1 ) * 100
```

What is the n, mean, median, standard deviation, and interquartile range of overall GPA?

```{r transform-descriptives}
mean(students$OVERALL_GPA, na.rm=TRUE)
median(students$OVERALL_GPA, na.rm=TRUE)
sd(students$OVERALL_GPA, na.rm=TRUE)
IQR(students$OVERALL_GPA, na.rm=TRUE)
summary(students$OVERALL_GPA, na.rm=TRUE)

library(psych)
describe(students$OVERALL_GPA)
```

What are the descriptives of overall GPA for each ethnicity group?

```{r transform-describeBy}
describeBy(students$OVERALL_GPA, group=students$ETHNICITY, mat=TRUE)
```

Aggregate the number of hours and hours attempted by each student.

```{r transform-aggregate}
hours <- aggregate(coursework$HOURS, by=list(coursework$ID), FUN=sum)
head(hours)
names(hours) <- c('ID','Hours')

hoursAttempted <- aggregate(coursework$HOURS_ATTEMPTED, by=list(coursework$ID), FUN=sum)
head(hoursAttempted)
names(hoursAttempted) <- c('ID', 'HoursAttempted')
```

Merge the `hours` and `hoursAttempted` to the students data frame.

```{r transform-merge}
students <- merge(students, hours, by='ID', all.x=TRUE)
students <- merge(students, hoursAttempted, by='ID', all.x=TRUE)
head(students)
```


# Visualize

```{r}
library(ggplot2)
```

Create a histogram of the number transfer credits students have earned.

```{r vis-histogram-transfer}
ggplot(students, aes(x=TRANSFER_CREDITS_EARNED)) + geom_histogram()
```

Change the bin with for each bar to 3 credits.

```{r vis-histogram-transfer2}
ggplot(students, aes(x=TRANSFER_CREDITS_EARNED)) + geom_histogram(binwidth=2)
```

Create a separate histogram for each school (i.e. `SCHOOL_CODE`).

```{r vis-histogram-transfer3}
ggplot(students, aes(x=TRANSFER_CREDITS_EARNED)) + geom_histogram() +
	facet_wrap(~ SCHOOL_CODE, ncol=1)
```

Create a density plot of the number of transfer credits.

```{r vis-histogram-transfer}
ggplot(students, aes(x=TRANSFER_CREDITS_EARNED)) + geom_density()
```

Create a bar plot ethnicity.

```{r vis-barplot-ethnicity}
ggplot(students, aes(x=ETHNICITY)) + geom_bar()
```

Color each bar based upon gender.

```{r vis-barplot-ethnicity}
ggplot(students, aes(x=ETHNICITY, fill=GENDER)) + geom_bar()
```

Try adding `position='dodge'` to `geom_bar`.

Color each bar based upon gender.

```{r vis-barplot-ethnicity}
ggplot(students, aes(x=ETHNICITY, fill=GENDER)) + geom_bar(position='dodge')
```

Now, create a separate bar plot for military versus civilian.

Color each bar based upon gender.

```{r vis-barplot-ethnicity}
ggplot(students, aes(x=ETHNICITY, fill=GENDER)) + geom_bar() + 
	facet_wrap(~ ACTIVE_MIL_STUDENT, ncol=1)
```

Plot the relationship between GPA and overall GPA.

```{r vis-scatterplot}
ggplot(students, aes(x=EC_GPA, y=OVERALL_GPA)) + geom_point()
```

Add a Loess regression line.

```{r vis-scatterplot2}
ggplot(students, aes(x=EC_GPA, y=OVERALL_GPA)) + geom_point() + geom_smooth()
```


# Model

Is there a difference in overall GPA between males and females?

```{r model-gendergpa}
t.test(OVERALL_GPA ~ GENDER, data=students)
```



