---
title: "Intro to R Workshop"
output: html_document
---

# Import

In the R chunk below, load in the `students.csv`, `coursework.csv`, and `demographics.csv` files into R.

```{r import-read}
#students <- 
#coursework <- 
#demographics <- 
```

## Viewing Data

Print the first six rows of `students` and the last six rows of `coursework`.

```{r import-head}
```

What is the `str`ucture of these data frames?

```{r import-str}
```

## Subsetting Data

Print all the students who live in Vermont.

```{r import-select}
```

Whare are all the different grade codes students have.

```{r import-unique}
```


# Tidy

Convert the appropriate date columns to actual dates (if they are not already).

```{r tidy-convertdates}
```

Convert `GENDER` to a factor.

```{r tidy-factor}
```

Convert `GRADE_CODE` to an ordered factor. You may wish to first convert all withdrawn grades (AW, MW, and W) to missing values first (i.e. `NA`).

```{r tidy-orderedfactor}
```

Convert any missing values for `TRANSFER_CREDITS_EARNED` to zero.

```{r tidy-missingtozero}
```


# Transform

What are the number of students for each ethnicity group?

```{r transform-table}
```

What is the proportion of students in each ethnicity group?

```{r transform-proptable}
```

What is the number of students in each ethnicity group by gender?

```{r transform-table2}
```

What is the proportion of students in each ethnicity group by gender where the rows equal 1 (or 100)?

```{r transform-proptable2}
```

What is the n, mean, median, standard deviation, and interquartile range of overall GPA?

```{r transform-descriptives}
```

What are the descriptives of overall GPA for each ethnicity group?

```{r transform-describeBy}
```

Aggregate the number of hours and hours attempted by each student.

```{r transform-aggregate}
```

Merge the `hours` and `hoursAttempted` to the students data frame.

```{r transform-merge}
```


# Visualize

```{r}
library(ggplot2)
```

Create a histogram of the number transfer credits students have earned.

```{r vis-histogram-transfer}
```

Change the bin with for each bar to 3 credits.

```{r vis-histogram-transfer2}
```

Create a separate histogram for each school (i.e. `SCHOOL_CODE`).

```{r vis-histogram-transfer3}
```

Create a density plot of the number of transfer credits.

```{r vis-histogram-transfer4}

```

Create a bar plot ethnicity.

```{r vis-barplot-ethnicity}

```

Color each bar based upon gender.

```{r vis-barplot-gender}

```

Try adding `position='dodge'` to `geom_bar`.

Color each bar based upon gender.

```{r vis-barplot-genderethnicity}

```

Now, create a separate bar plot for military versus civilian.

```{r vis-barplot-military}

```

Color each bar based upon gender.

```{r vis-barplot-gendercolor}

```

Plot the relationship between GPA and overall GPA.

```{r vis-scatterplot}

```

Add a Loess regression line.

```{r vis-scatterplot2}

```


# Model

Is there a difference in overall GPA between males and females?

```{r model-ttest}

```



