---
title: "Introduction"
author: "Jason Bryer and Kim Speerschneider"
output: html_document
---

# Why use R?


# Installation


# Navigating RStudio


# Packages

R packages are collections of functions, data, and documentation to extend the base functionality of R. As of March 2015 there are over 6,000 packages available on the [Comprehensive R Archive Network](http://cran.r-project.org) (CRAN), the main repository hosting R packages.


# Companion Package

The `irutils` package contains functions and data that will be used throughout this books. The package can be installed from [CRAN](http://cran.r-project.org):

```{r irutils, eval=FALSE}
install.packages('irutils')
```


# Conventions




# Colophon

This book was written using [R](http://www.r-project.org), [RStudio](http://www.rstudio.com), and [R Markdown](http://rmarkdown.rstudio.com/) on [Github](http://github.com/jbryer/IntroR). You can view and download the source code as well as contribute fixes and suggestions on the Github repository at [github.com/jbryer/IntroR](http://github.com/jbryer/IntroR).

```{r sessioninfo}
devtools::session_info()
```
